### 2020-04-18 dataclean.tex
more changes from last meeting

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Viewing and Cleaning the Data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Viewing and Cleaning the Data}
\label{section:dataclean}
\begin{figure*}[bt]
  \includegraphics[width=6.5in]{figs/sample_sn.pdf}
  \caption{GPU serial number view of life and failures. Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:gpuview}
\end{figure*}
As we need to recover durations in operation from this data, correct
processing involves time adjustments for switching between daylight
saving time and standard time and leap years. We perform this by
setting a reference time zone (Eastern time) and converting all
date-times from strings into POSIX date-time variables with the R
\pkg{lubridate} package \cite{lubridate}, which enables appropriate
date arithmetic and sensible date constructs for graphs.

@@

To get some intuition for the GPU lifetimes on Titan, we give two
views of 80 randomly selected GPUs (Figure~\ref{fig:gpuview}) and 80
randomly selected locations (Figure~\ref{fig:locview}).
\begin{figure*}[tb]
  \includegraphics[width=6.5in]{figs/sample_loc.pdf}
  \caption{GPU location view of life and failures.  Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:locview}
\end{figure*}
The GPU view
visually documents the life of the unit: when it was installed and
removed at various locations, its DBE and OTB events, and the last
time it was seen. The location view documents the life of a location:
when different GPUs were installed and removed, their OTB and DBE
events, and whether a removal was the last time the unit was seen on
the system. These views were critical to understanding the data and to
verifying various data processing decisions.

@@

The location view shows that each location was operational almost all
the time with small gaps when GPUs were changed out.

\fix{Were the GPUs swapped out as a blade or were individual GPUs
  swapped on a blade? Seems it was the full blade - this may further
  refine censoring. A DBE on one GPU triggers the replacement of the
  blade with the other three good gpus?}

>>>>

% !TEX root = main.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Viewing and Cleaning the Data
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Viewing and Cleaning the Data}
\label{section:dataclean}
\begin{figure*}[bt]
  \includegraphics[width=6.5in]{figs/sample_sn.pdf}
  \caption{Serial number view of GPU life and failures. Legend: black
    dots are installs, black lines are lifetimes at installed
    location, blue squares are OTB events, red triangles are DBE
    events, and black ] are ``last seen'' events.}
  \label{fig:gpuview}
\end{figure*}
As we need to recover durations in operation from this data, correct
processing involves time adjustments for switching between daylight
saving time and standard time and leap years. We perform this by
setting a reference time zone (Eastern time) and converting all
date-times from strings into POSIX date-time variables with the R
\pkg{lubridate} package \cite{lubridate}, which enables appropriate
date arithmetic and sensible date constructs for graphs.

@@

To get some intuition for the GPU lifetimes on Titan, we give two
views of 80 randomly selected GPUs (Figure~\ref{fig:gpuview}) and 80
randomly selected locations (Figure~\ref{fig:locview}).
\begin{figure*}[tb]
  \includegraphics[width=6.5in]{figs/sample_loc.pdf}
  \caption{Location view of GPU life and failures.  Legend: black dots
    are installs, black lines are lifetimes at installed location,
    blue squares are OTB events, red triangles are DBE events, and
    black ] are ``last seen'' events.}
  \label{fig:locview}
\end{figure*}
The GPU view
visually documents the life of each GPU unit: when it was installed and
removed at various locations, its DBE and OTB events, and the last
time it was seen. The location view documents the life of a location:
when different GPUs were installed and removed, their OTB and DBE
events, and whether a removal was the last time the unit was seen on
the system. These views were critical to understanding the data and to
verifying various data processing decisions.

@@

The location view shows that each location was operational almost all
the time with small gaps when GPUs were changed out. It is also
notable that OTB and DBE events are associated with a single GPU
although four GPUs are together on a blade. The proactive replacements
in the rework cycle were done by full blades. Events on single GPUs
were usually first swapped for a new blade, the failed GPU was
replaced on the blade, and the fixed blade then reused elsewhere. The
survival analysis of Sec.~\ref{section:survival} handles these nuances
by appropriate censoring.

### 2020-04-18 dataprep.tex
first changes reflecting meeting - more coming
more changes from last meeting

    \vspace{5em}
    \includegraphics[width=\columnwidth]{figs/TitanLayout.png}
  \end{minipage}
  \caption{Titan cabinet, blade, and floor layout. Note cooling
    airflow is bottom to top within cabinet.}
  \label{fig:layout}
\end{figure}

\fix{In the decomissioning video, I noticed there are actually 12+2+12
  cabinets, the middle two being different. Are the compute nodes the
  pairs of outside 12s and the two middle are service nodes?}

\fix{The logical 3d torus layout is $25\times 16\times 24$. How do
  the 25 and 16 dimensions map onto this? Are the 4 nodes actually very far
  apart in terms of the network layout via the two routers?}

\fix{Where does cool air arrive? Is it just dumped out at the top of
  the heat exchanger and circulate down? I noticed airholes in some
  floor tiles (more on the right than on the left). Is there cool air
  under the floor?}

\fix{How does the coolant circulate? I notice three sets of pipes on
  each side of each row: (2 x 3) x 8 = 48 sets.}

>>>>

    \vspace{5em}
    \includegraphics[width=\columnwidth]{figs/TitanLayout.png}
  \end{minipage}
  \caption{Titan cabinet, blade, and floor layout. Cooling airflow is
    bottom to top within a cabinet, taking ambient air in at the foot
    and relasing it back out at the top. Cooling fluid is passed
    through two heat exchangers, one below Cage 0 and one above Cage
    2.}
  \label{fig:layout}
\end{figure}

% \fix{Where does cool air arrive? Is it just dumped out at the top of
%   the heat exchanger and circulate down? I noticed airholes in some
%   floor tiles (more on the right than on the left). Is there cool air
%   under the floor?}

% \fix{How does the coolant circulate? I notice three sets of pipes on
%   each side of each row: (2 x 3) x 8 = 48 sets.}
